A Dependently Typed Multi-Stage Calculus \ifnotanonymous\thanks{Supported by organization x.}\fi

### 2019-06-14 main.tex
Split main.tex.

\section{Introduction}

\subsection{Multi-stage Programming}

% \subsubsection{多段階計算とは何か?}

Multi-stage programming enables us to generate and run fragments of code at runtime.
MetaOCaml\cite{oleg2014} provide features for multi-stage programming, which are brackets, escape, and run.
Brackets, written \verb|.< e >.| in MetaOCaml, makes a code value from \verb|e|.
Run, written \verb| run e | in MetaOCaml, runs a code value of \verb|e| and restore \verb|e|.
Escape, written \verb| .~ e | in MetaOCaml, expands a code value of \verb|e|.
Unlike run, escape is supposed to use in a fragment of code, that is, escape cannot be used as an alternative of run.
For example, the following MetaOCaml expression

\begin{verbatim}
        let plusone = .< fun x -> x + 1 >. in .< .~plusone 2 >.
\end{verbatim}
evaluates to \verb|.< (fun x -> x + 1) 2 >.|. We can get the result with run.
\begin{verbatim}
        run (let plusone = .< fun x -> x + 1 >. in .< .~plusone 2 >.)
\end{verbatim}
is reduced to 3.

% \subsubsection{CSPに関する補足説明}

Cross-stage Persistence (CSP) is another primitive of multi-stage programming, which enable us to embed computed values into a code value.
This is an example of CSP.
\begin{verbatim}
        let plusone = fun x -> x + 1 in
        let y = plusone 2 in  .< y >.
\end{verbatim}
This program evaluates to \verb|.< 3 >.| not  \verb|.< plusone 2 >.|.
This is because the variable \verb|y| is introduced into a code fragment using CSP.
Therefore, the variable \verb|y| is embedded after it was calculated.
CSP is very important when we write practical programs 
because CSP enables us to use library functions in code fragments as in \verb|.<List.combine [1;2] ['a';'b']>.|.

% \subsubsection{多段階計算のメリットをべき乗の例を用いて説明する}

The main application of multi-stage calculi is program optimization.
A famous example is power functions.
Usual power functions take two arguments, which are the base and the exponent.
We can make a specialized power function for given exponents with multi-stage programming and 
optimize them by unrolling a loop in functions for given exponents.
The optimized power function for the exponent of \verb|3| looks like \verb|power3 = fun x -> x * x * x|.
\verb|power3| is faster than ordinary \verb|power| function because it contains no loop.
Multi-stage programming can optimize functions which take more than two arguments 
by generating a code fragment optimized to a given argument.

% \subsubsection{ベクトル計算も高速化できる}

Multi-stage programming can also optimize vector calculation.
For example, \verb|vadd| function, which takes two vectors and return the sum of them,
is implemented with a loop in many cases.
We can unroll \verb|vadd| function for a given vector length and optimize it with multi-stage programming.
\red{MetaOCamlで書いた長さ付きベクトルの例?}

% \subsubsection{生成した行列計算のコードは特定のサイズに特化しており、他のサイズでは使えない。
% このため、間違って組み合わせると容易にエラーを引き起こす。既存の型システムではこのエラーを防止できない}

However, there is a sever problem in functions which are optimized by multi-stage programming.
Although unrolled \verb|vadd| function is optimized, we cannot use it for different vector length.
For example, when you optimize for the length of 5, you shouldn't use it for vectors of 3 lengths.
Otherwise, we will get a Segmentation Fault error.
This problem is serious but existing type systems for multi-stage calculi cannot prevent it.

% We call this problem \red{Specialized Code Problem}.

\subsection{Dependent Types}

% \subsubsection{依存型とは何か?}

Dependent types are types which is dependent on values.
We can use dependent types for securer programming.
For example, we can realize vectors with their sizes with them.
\begin{verbatim}
        Vector :: Int -> *
\end{verbatim}
\verb|Vector| is a type constor which takes the length of vectors, that is, 
\verb|Vector 3| is a type for vectors whose lengths are 3.
If \verb|vadd| function has the type of \verb|Vector n -> Vector n -> Vector n|,
we can confirm the two arguments of \verb|vadd| function has the same length.
\red{secure programming? finer grained typing? とか}

\subsection{Multi-stage Programming with Dependent Types}
\red{このsubsectionは短いので上とくっつけてもいいかもしれない}

% \subsubsection{「生成した行列計算のコードは特定のサイズに特化しており、他のサイズでは使えない。
% このため、間違って組み合わせると容易にエラーを引き起こす。」問題を依存型で解決する。}

As we pointed out in the above section,
functions optimized with multi-stage programming can take only restricted values as arguments.
When you optimize \verb|vadd| function for the length of 5, you should it only for 5 length vectors.
We introduce dependent types into a multi-stage calculus
so that the type system can guarantee optimized functions are used properly.

In this paper, we design new multi-stage calculus \LMD by 
merging dependent types into existing multi-stage calculus \LTP\cite{Hanada2014}.
\LMD is multi-stage calculus which contains dependent types.
\LMD makes multi-stage programming more safer with the power of dependent types.

\subsection{Organization of the Paper}

The organization of this paper is the following.
Section 2 gives an informal overview of \LMD.
Section 3 defines the syntax, type system, full reduction, evaluation contexts of \LMD.
Section 4 shows the properties of \LMD including Unique Decomposition.
Section 5 discusses future works.
Finally, Section 6 discusses related works.
\AI{Use $\backslash$label and $\backslash$ref!}
\AI{``Related work'' and ``future work'' (they are uncountable).}

\section{Informal Overview of \LMD}

We designed \LMD, a multi-stage calculus with dependent type.
\LMD is based on \LTP\cite{Hanada2014} by Hanada and Igarashi, which is a multi-stage calculus with CSP and
we introduced dependent types based on \LTP\cite{attapl}.
In this section, we check \LMD informally after checking \LTP and \LLF which are the basis of \LMD.

\red{\LTP\cite{Hanada2014}が頻発するが同一論文の複数回引用に何かルールはあるのか?}
\subsection{\LTP}

% quote and unquote

\LTP\cite{Hanada2014} is a multi-stage calculus with CSP by Hanada and Igarashi.
In \LTP, brackets and escape are written $\TB_\alpha M$ and $\TBL_\alpha M$, respectively.
The type of $\TB_\alpha M$ is $\TW_\alpha \tau$ if $M$ has type $\tau$.
The type of $\TBL_\alpha M$ is $\tau$ if $M$ has $\TW_\alpha \tau$.
% Please notice that if $\TBL_\alpha M$ is well-typed, $M$ has a code type from the typing rule of \LTP.
In addition to normal $\beta$-reduction, there is a reduction rule for brackets and escape.
\begin{align*}
	\TBL_\alpha (\TB_\alpha M) \longrightarrow M 
\end{align*}
It means escape cancels brackets.
% This reduction is called $\longrightarrow_\Lambda$ in \LMD.
% \red{この文はここに来るべきなのか? LTPの説明ではないが}

% transition variable and transition

The subscript $\alpha$ in $\TB_\alpha M$ is a \textit{transition variable} and
a sequence of transition variables is called \textit{transition}.
The empty transition is represented by $\epsilon$.
They are used to show the thickness of brackets.
For example, $\TB_\alpha (\lambda x:\I.x+10)$ is the fragment of code which becomes $(\lambda x:\I.x+10)$ after it was run once and
$\TB_\alpha \TB_\beta (\lambda x:\I.x+10)$, which is abbreviated as $\TB_{\alpha\beta} (\lambda x:\I.x+10)$,
becomes $(\lambda x:\I.x+10)$ after it was run it twice.
\red{run twiceは表現が微妙}

% transition abstraction and application

There are abstractions for transition variables and applications for transition abstraction in \LTP.
They look like $\Lambda\alpha.M$ and $M A$, respectively.
A transition abstraction binds a transition variable in a term.
For example, all $\alpha$ in $\Lambda\alpha.(\TB_\alpha (\lambda x:\I.x))$ are bound.
It is only natural there is a reduction rule for transition application in \LTP.
The rule is the following.
\begin{align*}
	(\Lambda\alpha.M)\ A \longrightarrow M[\alpha\mapsto A]
\end{align*}
For example, $\Lambda\alpha.(\TB_\alpha (\lambda x:\I.x))\ (\beta\gamma)$ reduces to $\TB_{\beta\gamma} (\lambda x:\I.x)$.

% transition-related symbols disappear when the empty transition is substituted.
% In short, about "run"

Another important rule about a transition variable is 
that symbols with transition variables disappear 
when the empty transition $\epsilon$ is substituted to the transition variable.
The purpose of this rule is express \verb|run| without any special symbol.
In multi-stage calculus, \verb|run| is a very important operator which changes quoted code to the original code.
For example, $(\TB_\epsilon (\lambda x:\I.x))$ is equivalent to $(\lambda x:\I.x)$.
Therefore, $\Lambda\alpha.(\TB_\alpha (\lambda x:\I.x))\ \epsilon$ becomes $\lambda x:\I.x$
In \LTP, \verb|run| is realized with application to the empty transition $\epsilon$.

% transition in typing rules

A type judgement of \LTP is of the form $\G \vdash M : \tau @ A$.
A transition in a judgement represents where the judgement is true.
For example, $\G \vdash (\lambda x:\I.x) : \I \to \I @ \alpha$ means 
term $\lambda x:\I.x$ has type $\I \to \I$ at transition $\alpha$.
Especially, terms without quoting exist at the empty transition $\epsilon$.
For example, $(\lambda x:\I.x)\ (1+2)$ is at $\epsilon$ transition and 
$\TB_\alpha (\lambda x:\I.x)$ is at $\alpha$ transition.
Terefore, transitions appear in typing rules, too.
\begin{center}
	\infrule{\G\vdash M:\tau @{A\alpha}}{\G\vdash \TB_{\alpha}M:\TW_{\alpha}\tau @A}{\TTB} \andalso
	\infrule{\G\vdash M:\TW_{\alpha}\tau @A}{\G\vdash \TBL_{\alpha}M:\tau @{A\alpha}}{\TTBL}
\end{center}
\TTB, corresponding to brackets, means 
if $M$ is typed $\tau$ at transition $A\alpha$ then $\TB_{\alpha}M$, quoted $M$, is typed $\TW_{\alpha}\tau$ at $A$.
\TTBL\ is converse of \TTB.

% CSP

CSP, cross-stage persistence, is an important feature of \LTP.
It enables us to embed value at an outer transition into an inner transition.
$\%$ is dedicated to CSP in \LTP.
For example, $\lambda a:\I.\Lambda\alpha.(\TB_\alpha (\lambda x:\I.x+\%_\alpha a))$
\red{inner / outer は適切か?}

% Omitting Residualization
% この段落は3章のM eのあとに、この制限の結果として...という形で入れる

% There is another important feature called program residualization in \LTP.
% It means that a generated code can be dumped into a file.
% We can load the dumped file and run it.
% The difficulty arises when program residualization is used with CSP.
% Transition variables are classified into two kinds in \LTP in order to deal with this difficulty.

\subsection{\LLF}

% \LLF
\LLF is a simple system of dependent types introduced in \cite{attapl}.
\red{system / type system / calculus?}
It is based on Edinburgh LF\cite{harper1993framework}.
Therefore, all constants and base types are declared in the signature.
The \LLF type theory generalizes simply typed lambda calculus
by replacing the function type $\tau\to\sigma$ with the dependent product type $\Pi x:\tau.\sigma$.

% Kind, Well-formed kind
In addition to ordinary typing rules like simply typed lambda calculus,
there are kinding rules, well-formed kinding rules, term equivalence rules, type equivalence rules, and kind equivalence rules in \LLF.
Kinding rules and well-formed kinding rules are 
introduced in order to prohibit making illegal types such as $\textrm{Vect}\ \textrm{Bool}$.
For a well-formed type $\tau$, $\G \vdash \tau :: K$ means that $\tau$ has a kind $K$ under the environment $\G$ and 
for a well-formed kind $K$, $\G \vdash K$ means that $K$ is a well-formed kind under an environment $\G$.

% Type Equality
Type equality rules are needed because the type equivalence is not obvious unlike simply typed lambda calculus.
For example, $\textrm{Vect}\ 7$ should be equivalent to $\textrm{Vect}\ (3+4)$
but they are not equivalent seemingly. Thus, we must define equivalence rules.
In \LLF, equivalence is expressed with a symbol of $\E$.
$\G \vdash M \E N$ means a term $M$ and a term $N$ are equivalent under the environment $\G$.
$\G \vdash \tau \E \sigma$ means a type $\tau$ and a type $\sigma$ are equivalent under the environment $\G$.
$\G \vdash K \E J$ means a kind $K$ and a kind $J$ are equivalent under the environment $\G$.

\subsection{Extending \LTP with Dependent Types}

Next, we develop \LMD by extending \LTP with \LLF-like dependent types.
From here, we use the word "stage" instead of "transition" 
because we develop a multi-stage calculus, not a multi-transition calculus.
\red{stageのほうが言葉としてふさわしいと言いたい}
There are three technical points in the extension from \LTP to \LMD.

\red{以下の3文をどこかに入れる必要がある}
\red{, which is called a \textit{stage variable} in \LMD.}
\red{ which is called a stage in \LMD.}
\red{Although we can apply any transition to a transition abstraction in \LTP,
	we restrict the transitions only to $\epsilon$ transition in \LMD to simplify the system.}

% Constants and Base Types

First, the way of handling of constants and base types is the difference between \LMD and \LTP.
We adopt a signature $\Sigma$ to handle constants and base types.
This is because a signature simplifies kinding rules relating to type variables.
A signature $\Sigma$ is composed of pairs of a base type and its kind or a constant and its type.
For example, if you want to use boolean, $\Sigma = \B::*, \text{true}:\B, \text{false}:\B$
\red{具体的な導出例を出したほうがよいか? また、なぜsimpleになるのを書くべきか?}

% Kidinding and Well-formed Kinding Rules

Second, we need kinding rules and well-formed kinding rules in order to extend \LMD with dependent types.
It was lucky that almost all rules are determined easily.
This is because multi-stage calculus and dependent types are almost orthogonal.
\red{orthogonal は抽象的すぎるか?}
Therefore, we can get kinding rules and well-formed kinding rules of \LMD just by 
attaching stage anotations to ones of \LLF.
For example, \KAbs-LF is a kinding rule for a dependent type in \LLF and \KAbs\ is a corresponding one.
\begin{center}
	\infrule{\G\vdash \tau :: * \andalso \G,x:\tau @A\vdash \sigma::J}{\G\vdash(\Pi x:\tau.\sigma) :: (\Pi x:\tau.J)}{\KAbs-LF} \\[2mm]
	\infrule{\G\V \tau :: *@A \andalso \G,x:\tau @A\V \sigma::J@A}{\G\V(\Pi x:\tau.\sigma) :: (\Pi x:\tau.J)@A}{\KAbs} \\[2mm]
\end{center}

% Equivalence Rules

Third, we also need type equivalence rules in \LMD as a dependent type system.
Although there are new primitives relating to stages which aren't in \LLF,
we can design all rules except \QPercent\ easily.
\red{design?}
\red{この段落が短い。後ろとくっつけてもいいが、段落の趣旨がボケる。}

% Q-% rule

The biggest problem in extension is handling the CSP symbol $\%_\alpha$ in \LMD.
The equivalence simple rule to handle CSP is \QCsp.
\begin{center}
	\infrule{\G\V M \E N : \tau @A}{\G\V\%_\alpha M \E \%_\alpha N : \tau @{A\alpha}}{\QCsp}
\end{center}
However, this rule isn't enough when the parameters of dependent types are cross-staed.
We will discuss this problem in a later section and solve this problem with the new rule \QPercent.
\begin{center}
	\infrule{\G\V M:\tau @{A\alpha} \andalso \G\V M:\tau @A}{\G\V\%_\alpha M \E M : \tau @{A\alpha}}{\QPercent}
\end{center}
\red{この段落でQ-Cspの持つ問題点を指摘すべきか? ただ、ここでmulmatの例を使うと3章と被る。}

\section{Formal Definition of \LMD}

In this section, we give a formal definition of \LMD, including
the syntax, full reduction, and type system.  In addition to full reduction,
in which any redex at any stage can be reduced, we also give staged reduction,
which models execution as programs.

\subsection{Syntax}

\AI{Rename ``type variables'' to ``type-level constants''. }
We assume the denumerable set of type variables, ranged over by
metavariables \(X, Y, Z\), the denumerable set of variables, ranged
over by \(x,y,z\), the denumerable set of constants, ranged over by
\(c\), and the denumerable set of stage variables, ranged over by
\(\alpha, \beta, \gamma\).  The metavariables \(A, B, C\) range over
sequences of stage variables; we write \(\varepsilon\) for the empty
sequence.  \AI{I prefer $\varepsilon$ for the nonempty sequence.}  \LMD is defined by the following grammar:

\AI{``Transition'' is still used.}

\begin{align*}
    % \textrm{Type variables}  &   &                          & X,Y,Z                                                                                                      \\
    % \textrm{Variables}       &   &                          & x,y,z                                                                                                      \\
    % \textrm{Stage variables} &   &                          & \alpha,\beta,\gamma                                                                                        \\
    % \textrm{Stage}           &   &                          & A,B,C                                                                                                      \\
    \textrm{Kinds}           &   & K,J,I,H,G                & ::= * \mid \Pi x:\tau.K                                                                                    \\
    \textrm{Types}           &   & \tau,\sigma,\rho,\pi,\xi & ::= X \mid \Pi x:\tau.\tau \mid \tau\ M \mid \TW_{\alpha} \tau \mid \F\alpha.\tau                          \\
%     \textrm{Constants}       &   &                          & c                                                                                                          \\
    \textrm{Terms}           &   & M,N,L,O,P                & ::= c \mid x \mid \lambda x:\tau.M\ \mid M\ M \mid \TB_\alpha M                                            \\
                             &   &                          & \ \ \ \ \mid \TBL_\alpha M \mid \Lambda\alpha.M \mid M\ \epsilon \mid \%_\alpha M                          \\  
    \textrm{Signature}       &   & \Sigma                   & ::= \emptyset \mid \Sigma, X::K \mid \Sigma, c:\tau                                                                        \\
    \textrm{Contexts}        &   & \Gamma                   & ::= \emptyset \mid  \Gamma,x:\tau @A\ (x\not\in\textrm{Domain}(\G)\ \text{and}\ x\not\in\textrm{FV}(\tau)) \\
\end{align*}

\AI{Add $M\,\alpha$ to terms.}

\AI{Is $O$ used as a metavariable?}
\AI{Domain and FV are not defined (yet).  I think we should introduce the notion of well-formed type environments by prose below and assume every type environment is well formed.}

% Description on meta variables

% Kinds

A kind, which is used to classify types, is either $*$, the kind of
proper types (types that terms inhabit), or $\Pi x\colon\tau.K$, the kind
of type operators that takes $x$ as an argument of type $\tau$ and returns a type
of kind $K$.
% Types
% of terms have kind $*$ and dependent types have $\Pi$-kinds.  For
% example, $\lambda x:\I.x$ has type $\Pi x:\I.\I$, which has $*$ kind.
% \red{この段落は短いので型の段落と結合するか?}
% Types
A type is a type variable $X$, which is declared in the signature with its kind, a dependent function type $\Pi x:\tau_1.\tau_2$,
an application of a type (operator) of $\Pi$-kind to a term, a code type $\TW_\alpha$, or an $\alpha$-closed type $\F\alpha.\tau$.
We sometimes abbreviate $\Pi x:\tau_1.\tau_2$ to $\tau_1 \rightarrow \tau_2$ if
$x$ is not a free variable in $\tau_2$. \AI{This should be mentioned after free variables are introduced.}
An application of a term to a dependent function type \AI{this sounds like $M\, (\Pi x:\tau_1.\tau_2)$} is, for example, $\text{Vector}\ 10$
if type variable $\text{Vector}$ has the kind $\Pi x:\I.*$.
A code type $\TW_\alpha \tau$ denotes a code fragment of a term of type $\tau$.
An $\alpha$-closed type corresponds to a runnable code fragment.
Type constructors tie strongly \AI{Is ``tie strongly'' standard?} in the order of 
an application of a term to a dependent function type $\tau\ M$,
a code type $\TW_{\alpha} \tau$,
a dependent function type $\Pi x:\tau.\tau$, 
an $\alpha$-closed type $\F\alpha.\tau$.
Therefore, $\F\alpha.\TW_{\alpha} \Pi x:\I.\text{Vector}\ 5$ is interpreted as
$\F\alpha.(\TW_{\alpha} (\Pi x:\I.(\text{Vector}\ 5)))$. \AI{I don't understand this rule...}

% Terms

Terms include ordinary (explicitly typed) \(\lambda\)-terms, constants,
whose type is declared in signature $\Sigma$, and the following five forms
related to multi-stage programming:
$\TB_\alpha M$ represents a code fragment; $\TB_\alpha M$ represents escape;
$\Lambda\alpha.M$ is a stage variable abstraction;
$M\ \epsilon$ is an application of $\epsilon$ to a stage variable \AI{I think the usage of ``application'' is wrong again}; and 
$\%_\alpha M$ is an operator for cross-stage persistence.
The prefix operators $\TB_\alpha, \TBL_\alpha$, and $\%_\alpha$ tie stronger than the two forms of applications
and applications are left-associative
and two abstractions extends as far to the right as possible.
For example, $\F\alpha.\lambda x:\I.\TB_\alpha x\ y$ means $\F\alpha.(\lambda x:\I.(\TB_\alpha x)\ y)$.

% Signature

We adopt a \LLF-like system, where constants and type variables for base types \AI{``for base types''?} are given in the signature $\Sigma$, which 
are a sequence of declarations of the form $c:\tau$ and $X:K$.
For example, when we use integers in \LMD, $\Sigma = \I :: *, 1:\I, 2:\I, \cdots$.  \AI{It sounds like we need an infinite signature...}
\AI{Change Int to Bool.}

% Free variables
As usual, the variable $x$ is bound in $\lambda x:\tau.M$
and the transition variable $\alpha$ is bound in $\Lambda \alpha.M$.
\AI{What about other binders such as $\Pi$?}
We identify $\alpha$-convertible terms and assume the names of bound variables are pairwise distinct.
We write $\FV(M)$ and $\FTV(M)$ for the set of free variables and the set of free stage variables in $M$, respectively.

% Contexts
Contexts \AI{Why not call them ``type environments'' to avoid confusion with evaluation contexts?} are sequences of triples of a variable, its type, and its stage.
$\textrm{Domain}$ is a function from a content to a set of variables defined in the context.
\AI{I'd prefer \textit{dom}.}
The side condition of $x\not\in\textrm{Domain}(\G)$ prevent multiple appearances of the same variable and
$x\not\in\textrm{FV}(\tau)$ is needed in the proof of Substitution Lemmas.

\AI{We should note that constants and type variables are not staged.}
\paragraph{Remark:}  \AI{Comparison with Hanada--Igarashi.}

\subsection{Reduction}

Next, we define full reduction for \LMD.
Before giving the definition of reduction, we define two kinds of substitutions.
Substitution $M[x\mapsto N]$ is the ordinary capture-avoiding substitution of
$N$ for $x$ in $M$, and we omit its definition here.
Substitution $M[\alpha \mapsto \epsilon]$ of the empty stage
for stage variable $\alpha$ in $M$ is defined below.
\AI{What about $\tau[x \mapsto N], K[x \mapsto N], \tau [\alpha \mapsto \epsilon]$ and $K[\alpha \mapsto \epsilon]$?}

\AI{Since we always substitute $\varepsilon$ for stage variables, we may want to introduce a special notation, something like $\textit{erase}_\alpha(M)$.}
\AI{We will use $[\alpha \mapsto \varepsilon]$ and $[\alpha \mapsto \beta]$.  The latter is renaming, so we can omit the definition.}

\begin{align*}
	% \newcommand{\SB}{}
	(\lambda x:\tau.M)[\alpha \mapsto \epsilon] & = \lambda x:(\tau[\alpha \mapsto \epsilon]).(M[\alpha \mapsto \epsilon])                                  \\
	(M\ N)[\alpha \mapsto \epsilon]             & = (M[\alpha \mapsto \epsilon])\ (N[\alpha \mapsto \epsilon])                                          \\
	(M\ \epsilon)[\alpha \mapsto \epsilon]      & = (M[\alpha \mapsto \epsilon])\ \epsilon                                                                \\
	(\TB_\alpha M)[\alpha \mapsto \epsilon]      & = M[\alpha \mapsto \epsilon]                                                                          \\
	(\TB_\beta M)[\alpha \mapsto \epsilon]      & = \TB_\beta (M[\alpha \mapsto \epsilon])                               & (\text{if } \alpha \neq \beta) \\
	(\TBL_\alpha M)[\alpha \mapsto \epsilon]     & =  M[\alpha \mapsto \epsilon]                                        &                               \\
	(\TBL_\beta M)[\alpha \mapsto \epsilon]     & = \TBL_\beta (M[\alpha \mapsto \epsilon])                              & (\text{if } \alpha \neq \beta) \\
	(\Lambda\beta.M)[\alpha \mapsto \epsilon]   & = \Lambda\beta.(M[\alpha \mapsto \epsilon])                            & (\text{if } \alpha \neq \beta) \\
	% (\Lambda\beta.M)[\alpha \mapsto \epsilon]   & = \Lambda\beta.M                                                     & (\text{if } \alpha = \beta)    \\
	(\%_\alpha M)[\alpha \mapsto \epsilon]       & = M[\alpha \mapsto \epsilon]                                         &                                  \\
	(\%_\beta M)[\alpha \mapsto \epsilon]       & = \%_{\beta[\alpha \mapsto \epsilon]}(M[\alpha \mapsto \epsilon])      & (\text{if } \alpha \neq \beta) \\
\end{align*}

\begin{definition}[Reduction]
  \AI{It's not clear what are defined here.  You are talking only about rules.
    Something like ``The relations $M \longrightarrow_\beta N$, ... are the least compatible relations closed under the rules below.''}
	There are three reduction rules ($\longrightarrow_\beta, \longrightarrow_\blacklozenge, \longrightarrow_\Lambda$) in \LMD.
	Congruence rules which are omitted from the definition.
	\begin{align*}
		 & (\lambda x:\tau.M) N \longrightarrow_\beta M[x \mapsto N]                       \\
		% & (\Pi x:\tau.\sigma) M \longrightarrow_\gamma \sigma[x \mapsto M] \\
		 & \TBL_\alpha \TB_\alpha M \longrightarrow_\blacklozenge M                        \\
		 & (\Lambda \alpha.M)\ \epsilon \longrightarrow_\Lambda M[\alpha \mapsto \epsilon]
	\end{align*}
	We write $ M \longrightarrow M'$ iff $ M \longrightarrow_\beta M'$, $ M \longrightarrow_\blacklozenge M'$, or $ M \longrightarrow_\Lambda M'$.
\end{definition}

The relation $\longrightarrow_\beta$ represents ordinary $\beta-$reduction in the \(\lambda\)-calculus; the relation 
$\longrightarrow_\blacklozenge$ represents that quotation $\TB_\alpha M$ is canceled by escape and $M$ is spliced into the code fragment surrounding the escape;
the relation $\longrightarrow_\Lambda$ means that a stage abstraction applied to the empty stage $\epsilon$ reduces to the body of abstraction
where $\epsilon$ is substituted for the stage variable.
Application of a non-$\epsilon$ stage to a stage abstraction is prohibited in order to simplify \LMD. \AI{This should have been mentioned earlier in the comparison with Hanada--Igarashi.}
There is no reduction rule for CSP as with Hanada and Igarashi \cite{Hanada2014}.
The CSP operator $\%_\alpha$ disappears when $\epsilon$ is substituted to $\alpha$.
% $\TB_\alpha$ and $\TBL_\alpha$ disappears in the same way as $\%_\alpha$.

\AI{A few examples are desirable.}

\subsection{Type System}

% 総論

In this section, we define the type system of \LMD.
% \LMD is little complicated because it contains dependent types.
It consists of six judgment forms for typing, kinding, well-formed kinding, term equivalence, type equivalence, and kind equivalence.
We list the judgments forms in Figure~\ref{fig:LMD-six-judgements}.
\AI{equality vs. equivalence?}
\AI{A few judgment forms do not match rules.}

\begin{figure}
	\begin{center}
		\begin{align*}
			\G & \V M : \tau @ A   & \text{typing} \\
			\G & \V \tau :: K @ A  & \text{kinding} \\
			\G & \V K \iskind @ A  & \text{well-formed kinding} \\
			\G & \V M \E N @ A     & \text{term equality} \\
			\G & \V \tau \E \sigma & \text{type equality} \\
			\G & \V K \E J @ A     & \text{kind equality}
		\end{align*}
		\caption{Six judgment forms of the type system of \LMD.}
		\label{fig:LMD-six-judgements}
	\end{center}
\end{figure}

\AI{Insert a tilde (which is a space without a line break) between ``Figure'' and ref.}
\begin{definition}[Typing]
  The typing relation $ \G \V M : \tau @ A $ is the least relation closed under the rules in Figure~\ref{fig:typing-rules}.
  \AI{Precisely speaking, it cannot be defined independently of other judgments.
  How is dependency between judgments?}
\end{definition}

\AI{``judgment'' vs. ``judgement''.}  We start from defining typing of
\LMD.  Typing rules of \LMD are shown in Figure
\ref{fig:typing-rules}.  The rule \TConst{} means any constants in the
signature can appear at any stage.  \AI{This should be discussed
  earlier.}  For example, if we have a signature $\Sigma$ which is
$\textrm{bool} :: *, \textrm{true}: \textrm{bool}, \textrm{false}:
\textrm{bool}$, the derivation tree in
Figure~\ref{fig:tconst-derivation-tree} is admissible.  \AI{Don't we
  need signature well formedness?  If $c:\tau \in \Sigma$, then $\tau$
  should be a well-formed type under ... what?}  The rules \TVar,
\TAbs, and \TApp{} are almost the same as those in the simply typed
lambda calculus.  \AI{... or \(\lambda\)LF?  We are working in a dependently type system...}  Additional conditions are that subterms must be
typed at the same stage (\TAbs{} and \TApp); the type
annotation/declaration on a variable has to be a proper type of kind
$*$ (\TAbs) at the stage where it is declared (\TVar{} and \TAbs).

\begin{figure}
	\begin{center}
		\begin{minipage}{0.50\hsize}
			\infer[\TConst]
			{\G \V \textrm{true}:\textrm{bool}@\alpha\beta}
			{\textrm{true}:\textrm{bool} \in \Sigma \andalso
				\ID{\G\V\textrm{bool}::*@\alpha\beta} \andalso
			}
			\caption{A derivation tree using \TConst}
			\label{fig:tconst-derivation-tree}
		\end{minipage}
	\end{center}
\end{figure}

% \TConv
\AI{Looks like a paper is written for those who know multi-stage calculi fairly well but don't know dependent type systems.  The reality is opposite.}
As in standard dependent type systems, \TConv{} allows us to replace
the type of a term with an equivalent one.  
In a type system which includes dependent types, this kind of rule is essential
because two types which have different shapes may be equivalent.
For example, when we use a vector type with its size ($\textrm{Vector}\ n$),
$\textrm{Vector}\ 5$ is equivalent to $\textrm{Vector}\ (4+1)$ obviously although they are not equivalent in apparently.

% Typing rules for a multi-stage calculus
\AI{This paragraph is not very informaticve.  We shouldn't just refer redears to previous work.}
The rules \TTB, \TTBL, \TGen, \TIns, and \TCsp{} are rules for a multi-stage calculus.
They are corresponding to brackets, escape, making a stage abstraction, 
application of $\epsilon$ stage to a stage abstraction, and cross-stage persistence, respectively.
Please check Hanada and Igarashi \cite{Hanada2014} and Tsukada and Igarashi \cite{Tsukada} for details of these rules.

\begin{figure}
	\begin{center}
		\infrule[\TConst]{c:\tau \in \Sigma \andalso \G\V \tau::*@A}{\G \V c:\tau @A} \\[2mm]
		\infrule[\TVar]{x:\tau @A \in \G \andalso \G\V \tau::*@A}{\G \V x:\tau @A} \\[2mm]
		\infrule[\TAbs]{\G\V \sigma::*@A\andalso\G,x:\sigma@A\V M:\tau @A}{\G\V(\lambda (x:\sigma).M):(\Pi (x:\sigma).\tau)@A} \\[2mm]
		\infrule[\TApp]{\G\V M:(\Pi (x:\sigma).\tau)@A \andalso \G\V N:\sigma@A}{\G\V M\ N : \tau[x\mapsto N]@A} \\[2mm]
		\infrule[\TConv]{\G\V M:\tau @A \andalso \G\V \tau\equiv \sigma :: K@A}{\G\V M:\sigma@A} \\[2mm]
		\infrule[\TTB]{\G\V M:\tau @{A\alpha}}{\G\V\TB_{\alpha}M:\TW_{\alpha}\tau @A} \andalso
		\infrule[\TTBL]{\G\V M:\TW_{\alpha}\tau @A}{\G\V\TBL_{\alpha}M:\tau @{A\alpha}} \\[2mm]
		\infrule[\TGen]{\G\V M:\tau @A \andalso \alpha\notin\rm{FTV}(\G)\cup\rm{FTV}(A)}{\G\V\Lambda\alpha.M:\forall\alpha.\tau @A} \\[2mm]
		\infrule[\TIns]{\G\V M:\forall\alpha.\tau @A}{\G\V M\ \epsilon:\tau[\alpha \mapsto \epsilon]@A} \andalso
		\infrule[\TCsp]{\G\V M:\tau @A}{\G\V \%_\alpha M:\tau @{A\alpha}}
		\caption{Typing Rules.}
		\label{fig:typing-rules}
	\end{center}
\end{figure}

\subsubsection{Type and Term Equivalence}

As usual in dependent type systems, type equivalence, used in \TConv,
is important.  The type equivalence judgment of the form
$\G \V \tau \E \sigma : K @ A$ means that types $\tau$ and $\sigma$
are equivalent as types of kind $K$ at stage $A$ under $\G$.
Figure~\ref{fig:type-equivalence-rules} shows the rules for type
equivalence.  Type equivalence is basically the least congruence
closed under term equivalence.  The rules for compatibility (closure
under type formation) are derived from corresponding typing rules in a
straightforward manner.  The rules are a little simpler than some
dependent type systems \AI{such as?}  because there is no abstraction
at the type level.

% \QCsp以外の説明
% We show type equivalence rules in Figure \ref{fig:type-equivalence-rules}.
% All rules except \QTRefl, \QTSym, \QTTrans, and \QTApp\ are generated naturally from the typing rules.
% \QTRefl, \QTSym, \QTTrans\ exist in order to make the type equivalence relation an equivalence relationship.
% The rule \QTApp\ means that if there are two equivalent $\Pi$ type and two equivalent terms,
% the results of applications are also equivalent.

>>>>

\input{introduction.tex}

\input{informal-overview.tex}

\input{formal-defnition.tex}

\input{properties.tex}

\input{related-work.tex}

\input{conclusions.tex}
